### 2021-05-08 article.Rmd
Update

---
title             : "Study sample characteristics in chronobiology and sleep research"
shorttitle        : "Study sample characteristics in chronobiology and sleep research"

author: 
  - name          : "Selma Tir"
    affiliation   : "1,2"
    email         : "selma.tir@ndcn.ox.ac.uk"
    role:
      - Data curation
      - Investigation
      - Software
      - Visualization
      - Writing – original draft
      - Writing – review & editing
  - name          : "Rhiannon White"
    affiliation   : "1,3"
    email         : "r.white.5@warwick.ac.uk"
    role:         
      - Investigation
      - Writing – original draft
      - Writing – review & editing
  - name          : "Manuel Spitschan"
    affiliation   : "1,4,5"
    corresponding : yes    # Define only one corresponding author
    email         : "manuel.spitschan@psy.ox.ac.uk"
    role:
      - Conceptualization
      - Data curation
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Software
      - Writing – original draft
      - Writing – review & editing

affiliation:
  - id            : "1"
    institution   : "Department of Experimental Psychology, University of Oxford"
  - id            : "2"
    institution   : "Sleep and Circadian Neuroscience Institute, Nuffield Department of Clinical Neurosciences, University of Oxford"
  - id            : "3"
    institution   : "Warwick Medical School, University of Warwick, United Kingdom"
  - id            : "4"
    institution   : "Centre for Chronobiology, Psychiatric Hospital of the University of Basel, Switzerland"
  - id            : "5"
    institution   : "Transfaculty Research Platform Molecular and Cognitive Neurosciences, University of Basel, Switzerland"

authornote: |
  Corresponding author: Manuel Spitschan (manuel.spitschan@psy.ox.ac.uk). This work ws funded by the Wellcome Trust (Sir Henry Wellcome Fellowship to MS 204686/Z/16/Z; Research Enrichment – Diversity & Inclusion WT 204686/Z/16/A) and Linacre College (Biomedical Sciences Junior Research Fellowship to MS).
abstract: |
  Many aspects of sleep and circadian physiology appear to be sensitive to participant-level characteristics. While recent research robustly highlights the importance of considering participant-level demographic information, it is not clear to what extent this information is available in the large body of already published literature. Here, we investigated study sample characteristics in the published sleep and chronobiology research over the past 40 years. 6,777 articles were identified and a random sample of 20% was included. The reporting of sample size, age, sex, gender, ethnicity, level of education, socio-economic status, and profession of the study population was scored, and any reported aggregate summary statistics for these variables were recorded. We found that while >90% of studies reported age or sex, all other variables were reported in <10% of cases. Sex balance greatly changed over the years, from a ~3:1 male to female ratio in the 1990s to a near-equal representation in the 2010s. We found that the majority of studies report at least sex or age, while other variables are typically not reported. Reporting quality is highly variable, indicating an opportunity to standardize reporting guidelines for participant-level characteristics to facilitate meta analyses.

keywords          : "demographics, ethnicity, sex, research participants, reporting, publishing, meta-science"
wordcount         : "X"

bibliography      : references.bib

floatsintext      : yes
figurelist        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "jou"
output            : papaja::apa6_pdf
---

```{r setup, include=FALSE}
# Version control
## R version 4.0.5 (2021-03-31)
## R Studio version 1.4.1106
## Microsoft Windows version 10.0.19041.329

# Load packages
library(papaja)
library(corrplot)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(ggExtra)
library(ggsci)
library(graphics)
library(gridExtra)
library(knitr)
library(latticeExtra)
library(mosaic)
library(pracma)
library(readr)
library(readxl)
library(reshape2)
library(stats) 
library(stringr)
library(tibble)
library(tidyr)
library(lemon)

r_refs("references_r.bib")
```

```{r Data, include=FALSE}
## Download data sheet from Github repository
## Change path if needed

# Set working directory
#setwd("~/Demographics Review")
setwd("~/Projects/sleep_circadian_demographics_data")

# Download scoring sheet 
data.original <- read_excel("data.xlsx")
```

# Introduction

  Many aspects of sleep and circadian physiology appear to be sensitive to characteristics of the studied population, most notably sex [@redline_effects_2004; @santhi_sex_2016; @cain_sex_2010; @anderson_sexual_2020; @mong_sleep_2011], age [@redline_effects_2004; @espiritu_aging-related_2008; @mander_sleep_2017; @desforges_sleep_1990; @li_sleep_2018; @bliwise_sleep_1993; @benloucif_responsiveness_2006; @duffy_aging_2015] and ethnicity [@ahn_scoping_2021; @goldstein_sleep_2020; @eastman_blacks_2012; @eastman_circadian_2016]. More generally, there is a large literature on individual differences on sleep and circadian physiology [@van_dongen_individual_2005; @kerkhof_inter-individual_1985; @tankova_circadian_1994; @horne_individual_1977; @baehr_individual_2000; @burgess_individual_2008; @santhi_spectral_2012; @chellappa_individual_2021], demonstrating the need to consider participant-level data.

The extent to which a scientific field's findings are generalisable depend very much on the representativeness of a given study sample. As an example, a recent study reviewed the reporting and analysis of sex in biological sciences research [@woitowich_10-year_2020]. The authors found that while sex inclusion has significantly increased over the past 10 years [@beery_sex_2011], sex-based analysis has not improved, despite recent policies and funder mandates [@clayton_policy_2014].

While research findings converge on participant-level demographic characteristics affecting outcomes, it is not clear to what extent this information is available in the large body of already published literature, nor to what extent it is even reported. Here, we address the question of participant-level demographic characteristics (age, sex, gender, ethnicity, level of education, socio-economic status, and profession of the study population) and reporting thereof in chronobiology and sleep research. Here, we extracted the study sample characteristics in a total of 1356 randomly sampled publications across the 8 top (ranked by the Journal Impact Factor) chronobiology and sleep research and subjected them to a comprehensive analysis.

# Methods

*Procedure*. Journal articles published between 1979 and 2019 (odd years) in the top eight sleep and chronobiology journals were considered. The listing of possible of target journals was based on a previously established list of journals implementing a hybrid strategy by consulting the Web of Science Master Journal List, domain-relevant expertise in sleep and chronobiology and consulting with a senior researcher with >25 years of experience in the field [@spitschan_transparency_2020]. From this previously derived list, we selected eight journals were selected based on their five-year Impact Factor, and included *Journal of Pineal Research* (ISSN: 0742-3098 / 1600-079X; 2018 5-year IF: 12.197), *Sleep*  (0161-8105 / 1550-9109; 5.588), *Journal of Sleep Research* (0962-1105 / 1365-2869; 3.951), *Sleep Medicine* (1389-9457 / 1878-5506; 3.934), *Journal of Clinical Sleep Medicine* (1550-9389 / 1550-9397; 3.855), *Journal of Biological Rhythms* (0748-7304 / 1552-4531; 3.349), Behavioral Sleep Medicine (1540-2002 / 1540-2010; 3.162), and *Chronobiology International* (0742-0528 / 1525-6073; 2.998). While *Sleep Medicine Reviews* also features in the list of journals, we did not include it as it primarily publishes reviews.

*Article inclusion*. 6,777 articles were identified through a MEDLINE search by the journal and including odd years. A random sample of 20% was initially selected for screening. Inclusion requirements included conducting original research in the English language, reporting human data, and recruiting volunteers. As such, animal studies, bibliographies, case reports, comments, conference proceedings, editorials, guidelines, letters, retracted publications, reviews, errata and corrigenda were excluded.

*Review and article extraction*. All included articles were reviewed for eligibility and coded by RW. The reporting of sample size, age, sex, gender, ethnicity, level of education, socioeconomic status, and profession of the study population was scored binarily (0 = not reported, 1 = reported), and any reported aggregate summary statistics for these variables were recorded. Funding source, geographical location and clinical focus of the article were examined, as well as whether data were analyzed by including any of the demographic variables as covariates.

```{r Inclusion & Exclusion, echo=FALSE, message=FALSE, warning=FALSE}
# Filter reviewed articles
data.original <- data.original %>%
  filter(article_fulltext_available_yn == 1)

# Compute total number of reviewed articles
n.original <- as.numeric(nrow(data.original))

# Separate included and excluded articles
data.included <- data.original %>%
  filter(article_included_yn == 1)

data.excluded <- data.original %>%
  filter(article_included_yn == 0)

# Check number of included and excluded articles is equal to total (return TRUE if yes)
check.inclusion <- (as.numeric(nrow(data.included)) + as.numeric(nrow(data.excluded))) == as.numeric(nrow(data.original))

# Compute percentage of included vs excluded articles, round to two decimals
percent.included <- round(as.numeric(nrow(data.included))/as.numeric(nrow(data.original)) * 100, 2)
percent.excluded <- round(as.numeric(nrow(data.excluded))/as.numeric(nrow(data.original)) * 100, 2)

# Stacked bar chart of included articles by journal and year of publication
figure1 <- ggplot(data.included, aes(x=article_year)) +
  geom_bar(stat="count", aes(fill=article_journal)) +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Number of included articles") +
  labs(fill="Journal") +
  ggtitle("Included and analysed articles by journal and year") +
  scale_x_continuous(breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_viridis_d() +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.title=element_text(face="bold", size=12), legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10))

# Compute percentage of articles excluded for other reasons
data.excluded.other <- data.excluded %>%
  filter(!is.na(article_excluded_reason_other))
percent.excluded.other <- round(nrow(data.excluded.other)/nrow(data.excluded) * 100, 2)

# Compute percentage of excluded articles by reason for exclusion
data.eligibility <- data.original
data.eligibility$article_excluded_reason <- tolower(data.eligibility$article_excluded_reason) 
data.eligibility <- data.eligibility %>%
  mutate(eligibility=if_else(!is.na(article_excluded_reason_other), "Not included for other reasons", 
                          if_else(article_included_yn == 1, "Included",
                                  if_else(article_excluded_reason == "not recruiting volunteers", 
                                          "Not recruiting volunteers",
                                          if_else(article_excluded_reason == "not an original research article",
                                                  "Not an original research article",
                                                  if_else(article_excluded_reason == "not reporting human data",
                                                          "Not reporting human data",
                                                          "NA"))))))

data.eligibility <- data.eligibility %>%
  count(eligibility, article_year) %>%
  ungroup()

data.eligibility <- data.eligibility %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(eligibility) %>%
  mutate(per=round(100*n/total,3))

data.eligibility$article_year <- as.numeric(data.eligibility$article_year)
data.eligibility$per <- as.numeric(data.eligibility$per)

# Stacked bar chart of excluded articles by reason for exclusion and year of publication
figure2 <- ggplot(data.eligibility, aes(x=article_year, y=per)) +
  geom_col(aes(fill=factor(eligibility, levels=c("Not an original research article",
                                                 "Not reporting human data",
                                                 "Not included for other reasons",
                                                 "Not recruiting volunteers",
                                                 "Included")))) +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  labs(fill="Eligibility") +
  ggtitle("Inclusion and reasons for exclusion") +
  scale_x_continuous(breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(labels=c("0%","25%","50%","75%","100%"), expand=c(0,0)) +
  scale_fill_manual(values=c("#29AF7FFF","#2D708EFF","#D8DE29FF","#482677FF")) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.title=element_text(face="bold", size=12), legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10))
```

  Data were coded in an Excel Spreadsheet and analyzed in R Studio (version 4.0.5). Reporting of funding, geographical location, and number of sub-studies for each article were investigated for the sample of articles that passed all eligibility criteria.

*Pre-registration*. We pre-registered our protocol (specified using the PRISMA-P template [@shamseer_preferred_2015; @prisma-p_group_preferred_2015]) on the Open Science Framework (https://osf.io/cu3we/).

*Materials, data and code availability*. All data underlying this manuscript are available on a public GitHub repository (https://github.com/hcvnl/sleep_circadian_demographics_data). The article was written in R [@R-base] using RMarkdown and papaja [@R-papaja], employing a series of additional R packages [@R-corrplot2017; @R-dplyr; @R-ggExtra; @R-ggformula; @R-ggplot2; @R-ggridges; @R-ggsci; @R-ggstance; @R-ggthemes; @R-gridExtra; @R-knitr; @R-lattice; @R-latticeExtra; @R-lemon; @R-Matrix; @R-mosaic; @R-mosaicData; @R-pracma; @R-readr; @R-readxl; @R-reshape2; @R-stringr; @R-tibble; @R-tidyr; @R-BayesFactor] and is fully reproducible.

```{r Funding, include=FALSE}
# Select articles that report funding
data.funding.name <- data.included %>%
  filter(funding_reported_yn == 1)

# Compute percentage of articles that report funding
percent.funding.name <- round(nrow(data.funding.name)/as.numeric(nrow(data.included)) * 100, 2)

# Select articles that report funding number
data.funding.num <- data.included %>%
  filter(funding_refnum_yn == 1)

# Compute percentage of articles that report funding number, out of those that report funding
percent.funding.num <- round(nrow(data.funding.num)/as.numeric(nrow(data.funding.name)) * 100, 2)

# Compute whether articles report funding and funding number
data.funding <- data.included %>%
  mutate(funding=if_else(!is.na(funding_names) & funding_refnum_yn==0, 
                         "Funding reported", 
                         if_else(!is.na(funding_names) & funding_refnum_yn==1, 
                                 "Funding and grant reported", 
                                 "Funding not reported")))

data.funding$funding <- as.factor(data.funding$funding)

# Count number of articles for funding condition by year of publication
data.funding <- data.funding %>%
  count(funding, article_year) %>%
  ungroup()

# Compute proportion of articles per sex category by year of publication
data.funding <- data.funding %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(funding, add=TRUE) %>%
  mutate(per=round(100*n/total,3))
data.funding$article_year <- as.numeric(data.funding$article_year)
data.funding$per <- as.numeric(data.funding$per)

# Stacked bar chart of reporting of funding and funding number by year of publication
figure3 <- ggplot(data.funding, aes(x=article_year, y=per)) +
  geom_col(aes(fill=factor(funding, levels=c("Funding not reported",
                                            "Funding reported",
                                            "Funding and grant reported")))) +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  labs(fill="") +
  scale_fill_manual(values=c("#482677FF","#DCE319FF","#FDE725FF")) +
  scale_x_continuous(breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(labels=c("0%","25%","50%","75%","100%"), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.margin=margin(0,0,0,0), legend.box.margin=margin(-10,0,-10,-10))

# Change funding names to lower case
data.funding.name$funding_names <- tolower(data.funding.name$funding_names)

# Create list of funding sources
list.funding <- data.funding.name$funding_names

# Split combinations of funding sources (multiple funding names)
# which are delimited by commas
list.funding <- as.list(strsplit(list.funding, ","))
list.funding <- unlist(list.funding)

# Count number of occurrences for each funding name
table.list.funding <- as.data.frame(table(list.funding))

# Combine funding names that fall under the same agency
table.list.funding$list.funding <- as.character(table.list.funding$list.funding)

# NIH (USA)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("nih", table.list.funding$list.funding)== TRUE | grepl("national institute", table.list.funding$list.funding) == TRUE | grepl("nimh", table.list.funding$list.funding)== TRUE | grepl("nhlbi", table.list.funding$list.funding)== TRUE | grepl("national heart lung", table.list.funding$list.funding)== TRUE | grepl("national cancer institute", table.list.funding$list.funding)== TRUE | grepl("nia", table.list.funding$list.funding)== TRUE | grepl("ninds", table.list.funding$list.funding)== TRUE | grepl("ninr", table.list.funding$list.funding)== TRUE | grepl("ninda", table.list.funding$list.funding)== TRUE, 
                              "NI", table.list.funding$list.funding))
# NHMRC (AUS)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("nhmrc", table.list.funding$list.funding)== TRUE | grepl("national health and medical research council", table.list.funding$list.funding)== TRUE,
                              "NHMRC", table.list.funding$list.funding))

# Wellcome Trust (GBR)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("wellcome", table.list.funding$list.funding)== TRUE,
                              "Wellcome Trust", table.list.funding$list.funding))

# CIHR (CAN)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("canadian institute", table.list.funding$list.funding)== TRUE,
                              "CIHR", table.list.funding$list.funding))

# NSFC (CHN)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("national natural science", table.list.funding$list.funding)== TRUE,
                              "NSFC", table.list.funding$list.funding))

# Academy of Finland (FIN)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("academy of finland", table.list.funding$list.funding)== TRUE,
                              "Academy of Finland", table.list.funding$list.funding))

# Inserm (FRA)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("inserm", table.list.funding$list.funding)== TRUE,
                              "Inserm", table.list.funding$list.funding))

# SNSF (CHE)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("swiss national", table.list.funding$list.funding)== TRUE,
                              "SNSF", table.list.funding$list.funding))

# Department of Veterans Affairs (USA)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("veterans affairs", table.list.funding$list.funding)== TRUE,
                              "Dept Vet Affairs", table.list.funding$list.funding))

# European Commission
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("european commission", table.list.funding$list.funding)== TRUE,
                              "EC", table.list.funding$list.funding))

# Group by funding agency
table.list.funding <- table.list.funding %>%
  group_by(list.funding) %>%
  summarise(N = sum(Freq))

# Compute percentage of articles that report funding from the National Institutes in the US
data.funding.NI <- data.funding.name %>%
  filter(grepl("nih", funding_names) | grepl("national institute", funding_names) | grepl("nimh", funding_names) | grepl("nhlbi", funding_names) | grepl("national heart lung", funding_names) | grepl("national cancer institute", funding_names) | grepl("nia", funding_names) | grepl("ninr", funding_names) | grepl("ninda", funding_names) | grepl("ninds", funding_names))

percent.funding.NI <- round(nrow(data.funding.NI)/as.numeric(nrow(data.included)) * 100, 2)

# Select articles that report funding number and funding from National Institutes in the US
data.funding.NI.num <- data.funding.NI %>%
  filter(funding_refnum_yn == 1)

percent.funding.NI.num <- round(nrow(data.funding.NI.num)/as.numeric(nrow(data.funding.NI)) * 100, 2)
```

```{r Study Country, echo=FALSE, message=FALSE, warning=FALSE}
# Set country code to upper case
data.included$study_country_values <- toupper(data.included$study_country_values)

# Separate articles for which country is reported explicitly vs inferred
data.country.reported <- data.included %>%
  filter(study_country_reported_explicitly_yn == 1)

data.country.inferred <- data.included %>%
  filter(study_country_reported_inferred_yn == 1)

# Compute percentage of articles where country is reported explicitly vs inferred
percent.country.reported <- round(nrow(data.country.reported)/nrow(data.included) * 100, 2)
percent.country.inferred <- round(nrow(data.country.inferred)/nrow(data.included) * 100, 2)

# Separate articles that report one or multiple countries of study
# Compute percentage of articles per study country
country.single <- data.included %>%
  filter(!grepl(",", study_country_values)) %>%
  group_by(study_country_values) %>%
  tally() %>%
  mutate(n=round((n/as.numeric(nrow(data.included))*100),1))

country.multiple <- data.included %>%
  filter(grepl(",", study_country_values)) %>%
  group_by(study_country_values) %>%
  tally() %>%
  mutate(n=round((n/as.numeric(nrow(data.included))*100),1))

# Compute percentage of articles where multiple countries of study are reported
percent.country.multiple <- round(nrow(country.multiple)/nrow(data.included) * 100, 2)

# Compute number of articles for the 8 most represented countries by year of publication
# Check in dataset "country.single" which countries are the 8 most represented

# Select articles where a single country is reported or inferred
data.country <- data.included %>%
  filter(!grepl(",", study_country_values)) %>%
  count(study_country_values, article_year) %>%
  ungroup()

# Change all countries that are not the 8 most represented countries to 'Other'
data.country$study_country_values[data.country$study_country_values != "AUS" &
                                    data.country$study_country_values != "CAN" &
                                    data.country$study_country_values != "DEU" &
                                    data.country$study_country_values != "FRA" &
                                    data.country$study_country_values != "GBR" &
                                    data.country$study_country_values != "USA" &
                                    data.country$study_country_values != "ITA" &
                                    data.country$study_country_values != "JPN"] <- "Other"

# Sum rows with 'Other' by year
data.country <- aggregate(cbind(n) ~ study_country_values + article_year, data=data.country, FUN=sum)

# Compute proportion of articles per country by year of publication
data.country <- data.country %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(study_country_values) %>%
  mutate(per=round(100*n/total,3))
data.country$article_year <- as.numeric(data.country$article_year)
data.country$per <- as.numeric(data.country$per)

# Stacked bar chart of 8 most represented study countries by year of publication
figure4 <- ggplot(data.country, aes(x=article_year, y=per)) +
  geom_col(aes(fill=factor(study_country_values, 
                           levels=c("Other","AUS","CAN","DEU","FRA",
                                    "GBR","ITA","JPN","USA")))) +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  labs(fill="Country") +
  scale_x_continuous(breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(labels=c("0%","25%","50%","75%","100%"), expand=c(0,0)) +
  scale_fill_viridis_d(direction=-1) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.title=element_text(face="bold", size=12), legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10))
```

```{r Number of sub-studies, echo=FALSE, message=FALSE, warning=FALSE}
# Compute percentage of articles per number of sub-studies
num.study <- data.included %>%
  group_by(num_substudies) %>%
  tally() %>%
  mutate(num_substudies=as.factor(num_substudies),
         n=round((n/as.numeric(nrow(data.included))*100),1))
percent.study.single <- num.study$n[1]
percent.study.multiple <- 1 - num.study$n[1]

### The rest  of the analysis will be conducted on included single studies
# Filter single studies
data <- data.included %>%
  filter(num_substudies == 1)
```

```{r Sample size, echo=FALSE, message=FALSE, warning=FALSE}
# Compute percentage of studies that report sample size
data.ss <- data %>%
  filter(sample_size_bycategory_reported_yn == 1)
percent.ss <- round(nrow(data.ss)/nrow(data) * 100, 2)

# Violin plot of sample size by year of publication with log10 scaling
data.ss$article_year <- as.factor(data.ss$article_year)
data.ss$sample_size_total_reported_values <- as.numeric(data.ss$sample_size_total_reported_values)

figure5 <- ggplot(data.ss, aes(x=article_year, y=sample_size_total_reported_values)) +
  geom_violin(fill="#440154FF") +
  theme_apa() +
  scale_y_continuous(trans='log10', breaks=c(10,1000,100000), 
                     labels=c("10","1,000","100,000")) +
  scale_x_discrete(breaks=c("1979","1989","1999","2009","2019")) +
  xlab("Year of publication") +
  ylab("Sample size, log-10 scaling") +
  annotation_logticks(sides="l") +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8))
```

```{r Age, message=FALSE, warning=FALSE, include=FALSE}
# Select articles reporting age
data.age <- data %>%
  filter(age_reported_mean_yn == 1 | age_reported_median_yn == 1 | age_reported_min_yn | age_reported_max_yn | age_reported_other_yn == 1 | age_reported_by_sex_yn == 1)

# Compute percentage of articles reporting age
percent.age <- round(nrow(data.age)/nrow(data) * 100, 2)

## Correlate reporting of age variables by year
# select year and age columns
data.age.year <- data.age %>%
  select(article_year, age_reported_mean_yn, age_reported_sd_yn, age_reported_median_yn, 
         age_reported_min_yn, age_reported_max_yn, age_reported_iqr_yn, age_reported_other_yn)

# Replace 0 by -1
data.age.year$age_reported_mean_yn[data.age.year$age_reported_mean_yn == 0] <- -1
data.age.year$age_reported_sd_yn[data.age.year$age_reported_sd_yn == 0] <- -1
data.age.year$age_reported_median_yn[data.age.year$age_reported_median_yn == 0] <- -1
data.age.year$age_reported_min_yn[data.age.year$age_reported_min_yn == 0] <- -1
data.age.year$age_reported_max_yn[data.age.year$age_reported_max_yn == 0] <- -1
data.age.year$age_reported_iqr_yn[data.age.year$age_reported_iqr_yn == 0] <- -1
data.age.year$age_reported_other_yn[data.age.year$age_reported_other_yn == 0] <- -1

# Replace NAs
data.age.year[is.na(data.age.year)] <- -1

# Group by year and compute average for each sex variable
age.corr.year <- data.age.year %>%
  group_by(article_year) %>%
  summarise(mean=mean(age_reported_mean_yn), 
            sd=mean(age_reported_sd_yn), median=mean(age_reported_median_yn),
            min=mean(age_reported_min_yn), max=mean(age_reported_max_yn),
            iqr=mean(age_reported_iqr_yn),
            other=mean(age_reported_other_yn))  %>%
  ungroup()

age.corr.year <- age.corr.year %>%
  group_by(article_year) %>%
  mutate(total=((mean+sd+median+min+max+iqr+other)/7))

# Reshape data
age.corr.year <- melt(age.corr.year,id.vars='article_year',
                      measure.vars=c('mean','sd','median','min',
                                     'max','iqr','other','total'))

# Heat map of reporting of age variables by year of publication
figure6 <- ggplot(data=age.corr.year, aes(x=variable, y=article_year, fill=value)) +
  geom_tile() +
  theme_apa() +
  ylab(label="Year of publication") +
  xlab("Age variable") +
  scale_y_continuous(breaks=c(1979,1989,1999,2009,2019), expand=c(0.02,0),
                     labels=c("                   1979", "1989","1999","2009","2019")) +
  scale_fill_viridis_c(end=0.6, direction=-1, name="Reporting",
                       limits=c(-1,1), labels=c("0.0","0.25","0.5","0.75","1")) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8), 
        legend.position="right", legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10), legend.title=element_text(face="bold"))

## Correlate reporting of age variables by journal
# Select journal and age columns
data.age.journal <- data.age %>%
  select(article_journal, age_reported_mean_yn, age_reported_sd_yn, age_reported_median_yn, 
         age_reported_min_yn, age_reported_max_yn, age_reported_iqr_yn, age_reported_other_yn)

# Replace 0 by -1
data.age.journal$age_reported_mean_yn[data.age.journal$age_reported_mean_yn == 0] <- -1
data.age.journal$age_reported_sd_yn[data.age.journal$age_reported_sd_yn == 0] <- -1
data.age.journal$age_reported_median_yn[data.age.journal$age_reported_median_yn == 0] <- -1
data.age.journal$age_reported_min_yn[data.age.journal$age_reported_min_yn == 0] <- -1
data.age.journal$age_reported_max_yn[data.age.journal$age_reported_max_yn == 0] <- -1
data.age.journal$age_reported_iqr_yn[data.age.journal$age_reported_iqr_yn == 0] <- -1
data.age.journal$age_reported_other_yn[data.age.journal$age_reported_other_yn == 0] <- -1

# Replace NAs
data.age.journal[is.na(data.age.journal)] <- -1

# Group by journal and compute average for each sex variable
age.corr.journal <- data.age.journal %>%
  group_by(article_journal) %>%
  summarise(mean=mean(age_reported_mean_yn), sd=mean(age_reported_sd_yn),
            median=mean(age_reported_median_yn),min=mean(age_reported_min_yn),
            max=mean(age_reported_max_yn), iqr=mean(age_reported_iqr_yn),
            other=mean(age_reported_other_yn)) %>%
  ungroup()

age.corr.journal <- age.corr.journal %>%
  group_by(article_journal) %>%
  mutate(total=((mean+sd+median+min+max+iqr+other)/7))

# Reshape data
age.corr.journal <- melt(age.corr.journal,id.vars='article_journal',
                         measure.vars=c('mean','sd','median','min',
                                        'max','iqr','other','total'))

# Heat map of reporting of age variables by journal of publication
figure7 <- ggplot(data=age.corr.journal, 
                   aes(x=variable, y=article_journal, fill=value)) +
  geom_tile() +
  theme_apa() +
  ylab(label="Journal of publication") +
  xlab("Age variable") +
  scale_fill_viridis_c(end=0.6, direction=-1, name="Reporting",
                       limits=c(-1,1), labels=c("0.0","0.25","0.5","0.75","1")) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.position="right", legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10), legend.title=element_text(face="bold"))

## Regression for age by year of publication
# filter articles that report mean age in years
data.age.mean <- data.age %>%
  filter(age_reported_mean_yn == 1) %>%
  filter(!is.na(as.numeric(age_reported_mean_value))) %>%
  select(article_pmid, article_year, age_reported_mean_value)

# filter articles that report mean age in weeks or months
data.age.mean.not.years <- data.age %>%
  filter(age_reported_mean_yn == 1) %>%
  filter(is.na(as.numeric(age_reported_mean_value))) %>%
  select(article_pmid, article_year, age_reported_mean_value)

# Change age from days / weeks / months to years, with 1 month = 4.3 weeks = 30.4 days
# age not in years is entered with "months" or "weeks"
data.age.mean.not.years <- data.age.mean.not.years %>%
  separate(age_reported_mean_value, 
           into=c("age_reported_mean_value", "text"),
           sep=" ")
data.age.mean.not.years$age_reported_mean_value <- as.numeric(data.age.mean.not.years$age_reported_mean_value)

data.age.mean.not.years <- data.age.mean.not.years %>%
  mutate(age_reported_mean_value=ifelse(text=="weeks", 
                                        age_reported_mean_value / 4.3 / 12,
                                        if_else(text=="days",
                                                age_reported_mean_value / 30.4 / 12,
                                                age_reported_mean_value / 12))) %>%
  select(article_pmid, article_year, age_reported_mean_value)# else case is in months

# Merge back mean age datasets
data.age.mean <- rbind(data.age.mean, data.age.mean.not.years)

data.age.mean$age_reported_mean_value <- as.numeric(data.age.mean$age_reported_mean_value)

# Check regression statistics
summary(lm(age_reported_mean_value ~ article_year, data=data.age.mean))

# Violin plot with regression line and equation for mean age by year of publication
figure8 <- ggplot(data=data.age.mean, aes(x=article_year, y=age_reported_mean_value)) +
  geom_smooth(method="lm", color="#440154FF") +
  geom_point(size=1, color="#33638DFF") +
  theme_apa() +
  ylab("Mean age (years)") +
  xlab("Year of publication") +
  scale_x_continuous(limits=c(1979,2019),breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(limits=c(0,80), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8))
```

```{r Sex, echo=FALSE, message=FALSE, warning=FALSE}
# Select sex variables       
data.sex <- data %>%
  select(article_year, sex_reported_yn,
         sex_reported_num_male, sex_reported_percent_male,
         sex_reported_num_female,sex_reported_percent_female)

# Compute whether males or females are reported at all
data.sex$male <- ifelse(is.na(data.sex$sex_reported_num_male) & is.na(data.sex$sex_reported_percent_male),
                             "No",
                             "Yes")
data.sex$male <- as.factor(data.sex$male)

data.sex$female <- ifelse(is.na(data.sex$sex_reported_num_female) & is.na(data.sex$sex_reported_percent_female),
                             "No",
                             "Yes")
data.sex$female <- as.factor(data.sex$female)

# Compute whether articles report males, females, both or unspecified
data.sex <- data.sex %>%
  mutate(report=if_else(male == "Yes" & female == "Yes", "Both Sexes", 
                          if_else(male == "Yes" & female == "No", "Male",
                                  if_else(male == "No" & female == "Yes", "Female", "Unspecified"))))

# Count number of articles for each sex condition by year of publication
data.sex <- data.sex %>%
  count(report, article_year) %>%
  ungroup()

# Compute proportion of articles per sex category by year of publication
data.sex <- data.sex %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(report, add=TRUE) %>%
  mutate(per=round(100*n/total,3))
data.sex$article_year <- as.numeric(data.sex$article_year)
data.sex$per <- as.numeric(data.sex$per)

# Stacked bar chart of sex reporting by year of publication
figure9 <- ggplot(data.sex, aes(x=article_year, y=per)) +
  geom_col(aes(fill=factor(report, levels=c("Unspecified",
                                            "Female",
                                            "Both Sexes",
                                            "Male")))) +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  labs(fill="Sex") +
  scale_fill_manual(values=c("#482677FF","#D8DE29FF","#29AF7FFF","#2D708EFF")) +
  scale_x_continuous(breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(labels=c("0%","25%","50%","75%","100%"), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.title=element_text(face="bold"),
        legend.margin=margin(0,0,0,0), legend.box.margin=margin(-10,0,-10,-10))

# Select articles for which sex is reported
data.sex.reported <- data %>%
  filter(sex_reported_yn == 1) %>%
  mutate(sample_size_total_reported_values=as.numeric(sample_size_total_reported_values),
         sex_reported_num_male=as.numeric(sex_reported_num_male),
         sex_reported_percent_male=as.numeric(sex_reported_percent_male),
         sex_reported_num_female=as.numeric(sex_reported_num_female),
         sex_reported_percent_female=as.numeric(sex_reported_percent_female))

# Select articles for which intersex is reported
data.intersex <- data %>%
  filter(sex_reported_intersex_yn == 1)

# compute percentage of articles where sex and intersex is reported
percent.sex <- round(nrow(data.sex.reported)/nrow(data) * 100, 2)
percent.intersex <- round(nrow(data.intersex)/nrow(data) * 100, 2)

# Compute whether males or females are recruited in the articles that report  sex
data.sex.reported$male <- ifelse(is.na(data.sex.reported$sex_reported_num_male) &
                                   is.na(data.sex.reported$sex_reported_percent_male),
                             "No",
                             "Yes")
data.sex.reported$male <- as.factor(data.sex.reported$male)

data.sex.reported$female <- ifelse(is.na(data.sex.reported$sex_reported_num_female) &
                                     is.na(data.sex.reported$sex_reported_percent_female),
                             "No",
                             "Yes")
data.sex.reported$female <- as.factor(data.sex.reported$female)

# Compute whether articles report males, females or both
data.sex.reported <- data.sex.reported %>%
  mutate(report=if_else(male == "Yes" & female == "Yes", "Both Sexes", 
                          if_else(male == "Yes" & female == "No", "Male",
                                  if_else(male == "No" & female == "Yes", "Female", "NA"))))

# Compute percentage of articles that only recruit males or females
data.sex.male <- data.sex.reported %>%
  filter(report=="Male")
percent.sex.male <- round(nrow(data.sex.male)/nrow(data.sex.reported) * 100, 2)

data.sex.female <- data.sex.reported %>%
  filter(report=="Female")
percent.sex.female <- round(nrow(data.sex.female)/nrow(data.sex.reported) * 100, 2)

# Compute percentage of articles reporting on a sex-dependent feature
data.sex.male.focus <- data.sex.male %>%
  filter(sample_focus_sex_dependent_yn == 1)
percent.sex.male.focus <- round(nrow(data.sex.male.focus)/nrow(data.sex.male) * 100, 2)

data.sex.female.focus <- data.sex.female %>%
  filter(sample_focus_sex_dependent_yn == 1)
percent.sex.female.focus <- round(nrow(data.sex.female.focus)/nrow(data.sex.female) * 100, 2)

# Compute percentage of studies reporting age by sex
data.sex.age <- data %>%
  filter(age_reported_by_sex_yn == 1)
percent.sex.age <- round(nrow(data.sex.age)/nrow(data) * 100, 2)
```

```{r Other demographics, echo=FALSE, message=FALSE, warning=FALSE}
# Select single studies and other demographic variables
data.demo <- data %>%
  select(article_year, ethnicity_reported_yn, 
         education_reported_mean_yn, education_reported_yn,
         profession_reported_yn, ses_reported_yn)

# Compute the reporting of the different demographic variables
data.demo <- data.demo %>%
  mutate(education=if_else(education_reported_mean_yn == 0 & education_reported_yn == 0, "No", "Yes"),
         ethnicity=if_else(ethnicity_reported_yn == 1, "Yes", "No"),
         profession=if_else(profession_reported_yn == 1, "Yes", "No"),
         ses=if_else(ses_reported_yn == 1, "Yes", "No"))

# Count number of articles for each demographic variable by year of publication
data.demo.education <- data.demo %>%
  count(education, article_year) %>%
  ungroup()
data.demo.ethnicity <- data.demo %>%
  count(ethnicity, article_year) %>%
  ungroup()
data.demo.profession <- data.demo %>%
  count(profession, article_year) %>%
  ungroup()
data.demo.ses <- data.demo %>%
  count(ses, article_year) %>%
  ungroup()

# Compute proportion of articles by year of publication
data.demo.education <- data.demo.education %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(education, add=TRUE) %>%
  mutate(per=round(100*n/total,3),
         variable="edu") %>%
  select(article_year, variable, education, per)
colnames(data.demo.education)[3] <- "report"

data.demo.ethnicity <- data.demo.ethnicity %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(ethnicity, add=TRUE) %>%
  mutate(per=round(100*n/total,3),
         variable="ethni") %>%
  select(article_year, variable, ethnicity, per)
colnames(data.demo.ethnicity)[3] <- "report"

data.demo.profession <- data.demo.profession %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(profession, add=TRUE) %>%
  mutate(per=round(100*n/total,3),
         variable="prof") %>%
  select(article_year, variable, profession, per)
colnames(data.demo.profession)[3] <- "report"

data.demo.ses <- data.demo.ses %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(ses, add=TRUE) %>%
  mutate(per=round(100*n/total,3),
         variable="ses") %>%
  select(article_year, variable, ses, per)
colnames(data.demo.ses)[3] <- "report"

# Merge back the datasets together
merge1 <- rbind(data.demo.education, data.demo.ethnicity)
merge2 <- rbind(merge1, data.demo.profession)
data.demo <- rbind(merge2, data.demo.ses)
data.demo$article_year <- as.numeric(data.demo$article_year)
data.demo$per <- as.numeric(data.demo$per)

# Scatter plot of reporting of other demographic variables by year of publication
data.demo.reported <- data.demo %>%
  filter(report == "No") %>%
  mutate(per = (100 - per))

figure10 <- ggplot(data.demo.reported, aes(x=article_year, y=per, color=factor(variable))) +
  geom_point(size=2.5) +
  geom_line(linetype="dashed") +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  scale_color_viridis_d(name="Variable reported", 
                        labels=c("Education","Ethnicity","Profession","Socio-economic status")) +
  scale_x_continuous(limits=c(1979,2019), breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(labels=c("0%","25%","50%","75%","100%"), limits=c(0,100)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.title=element_text(face="bold", size=12),
        legend.margin=margin(0,0,0,0), legend.box.margin=margin(-10,0,-10,-10))

# Compute percentage of articles that report education
data.education <- data %>%
  filter(education_reported_mean_yn == 1 | education_reported_yn == 1)
percent.education <- round(nrow(data.education)/nrow(data) * 100, 2)

# Compute distribution of the number of categories reported for education
data.education$education_reported_numcategory <- as.numeric(data.education$education_reported_numcategory)

figure11A <- ggplot(data=data.education, aes(x=education_reported_numcategory)) +
  geom_histogram(binwidth=1, aes(y=..density..), colour="black", fill="#2D708EFF") +
  geom_vline(aes(xintercept=mean(education_reported_numcategory, na.rm=TRUE)), 
             color="#D8DE29FF", size=1) +
  theme_apa() +
  ylab("Density") +
  xlab("Education") +
  scale_x_continuous(limits=c(0.5,8.5), breaks=c(1,2,3,4,5,6,7,8)) +
  scale_y_continuous(limits=c(0,0.5), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8))

# Compute percentage of articles that report ethnicity
data.ethnicity <- data %>%
  filter(ethnicity_reported_yn == 1)
percent.ethnicity <- round(nrow(data.ethnicity)/nrow(data) * 100, 2)

# Compute distribution of the number of categories reported for ethnicity
data.ethnicity$ethnicity_reported_numcategory <- as.numeric(data.ethnicity$ethnicity_reported_numcategory)
figure11B <- ggplot(data=data.ethnicity, aes(x=ethnicity_reported_numcategory)) +
  geom_histogram(binwidth=1, aes(y=..density..), colour="black", fill="#2D708EFF") +
  geom_vline(aes(xintercept=mean(ethnicity_reported_numcategory, na.rm=TRUE)), 
             color="#D8DE29FF", size=1) +
  theme_apa() +
  ylab("Density") +
  xlab("Ethnicity") +
  scale_x_continuous(limits=c(0.5,8.5), breaks=c(1,2,3,4,5,6,7,8)) +
  scale_y_continuous(limits=c(0,0.5), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8))

# Compute percentage of articles that report profession
data.profession <- data %>%
  filter(profession_reported_yn == 1)
percent.profession <- round(nrow(data.profession)/nrow(data) * 100, 2)

# Compute distribution of the number of categories reported for profession
data.profession$profession_reported_numcategory <- as.numeric(data.profession$profession_reported_numcategory)
figure11C <- ggplot(data=data.profession, aes(x=profession_reported_numcategory)) +
  geom_histogram(binwidth=1, aes(y=..density..), colour="black", fill="#2D708EFF") +
  geom_vline(aes(xintercept=mean(profession_reported_numcategory, na.rm=TRUE)), 
             color="#D8DE29FF", size=1) +
  theme_apa() +
  ylab("Density") +
  xlab("Profession") +
  scale_x_continuous(limits=c(0.5,8.5), breaks=c(1,2,3,4,5,6,7,8)) +
  scale_y_continuous(limits=c(0,0.5), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8))

# Compute percentage of articles that report ses
data.ses <- data %>%
  filter(ses_reported_yn ==1)
percent.ses <- round(nrow(data.ses)/nrow(data) * 100, 2)

# Compute distribution of the number of categories reported for ses
data.ses$ses_reported_numcategory <- as.numeric(data.ses$ses_reported_numcategory)
figure11D <- ggplot(data=data.ses, aes(x=ses_reported_numcategory)) +
  geom_histogram(binwidth=1, aes(y=..density..), colour="black", fill="#2D708EFF") +
  geom_vline(aes(xintercept=mean(ses_reported_numcategory, na.rm=TRUE)), 
             color="#D8DE29FF", size=1) +
  theme_apa() +
  ylab("Density") +
  xlab("SES") +
  scale_x_continuous(limits=c(0.5,8.5), breaks=c(1,2,3,4,5,6,7,8)) +
  scale_y_continuous(limits=c(0,0.5), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8)) 
```

```{r Focus, echo=FALSE, message=FALSE, warning=FALSE}
## Sex feature
# Compute percentage of articles focusing on a sex dependent feature
data.focus.sex <- data %>%
  filter(sample_focus_sex_dependent_yn == 1)
percent.focus.sex <- round(nrow(data.focus.sex)/nrow(data) * 100, 2)

# Compute number of articles per sex dependent feature 
sex.dependent <- data.focus.sex %>%
  group_by(sample_focus_sex_dependent) %>%
  tally() %>%
  mutate(n=round((n/nrow(data.focus.sex)*100),1))

# Pie chart of the reported sex dependent features 
pc.sex.dependent <- ggplot(data=sex.dependent, aes(x="", y=n, fill=sample_focus_sex_dependent)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  geom_text(aes(label=paste0(n, "%")), position=position_stack(vjust=0.5)) +
  ggtitle("Articles focused on a sex dependent feature") +
  labs(fill="Sex dependent feature") +
  scale_fill_brewer(palette="RdBu") +
  theme_apa()

## Clinical feature
# Compute percentage of articles focusing on a clinical feature
data.focus.clinical <- data %>%
  filter(sample_focus_clinical_yn == 1)
percent.focus.clinical <- round(nrow(data.focus.clinical)/nrow(data) * 100, 2)

# Compute number of articles per clinical feature 
clinical <- data.focus.clinical %>%
  group_by(sample_focus_clinical) %>%
  tally() %>%
  mutate(n=round((n/nrow(data.focus.clinical)*100),1))

# Pie chart of the reported clinical features 
pc.clinical <- ggplot(data=clinical, aes(x="", y=n, fill=sample_focus_clinical)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  geom_text(aes(label=paste0(n, "%")), position=position_stack(vjust=0.5)) +
  ggtitle("Articles focused on a clinical feature") +
  labs(fill="Clinical feature") +
  scale_fill_brewer(palette="RdBu") +
  theme_void()

## Twins
# Compute percentage of articles focusing on twins
data.focus.twins <- data %>%
  filter(sample_focus_twins == 1)
percent.focus.twins <- round(nrow(data.focus.twins)/nrow(data) * 100, 2)

## University students
# Compute percentage of articles focusing on university students
data.focus.students <- data %>%
  filter(sample_focus_university_students == 1)
percent.focus.students <- round(nrow(data.focus.students)/nrow(data) * 100, 2)

## Shift workers
# Compute percentage of articles focusing on shift workers
data.focus.shift <- data %>%
  filter(sample_focus_shift_workers == 1)
percent.focus.shift <- round(nrow(data.focus.shift)/nrow(data) * 100, 2)

## Pregnant women
# Compute percentage of articles focusing on pregnant women
data.focus.pregnant <- data %>%
  filter(sample_focus_pregnant_women == 1)
percent.focus.pregnant <- round(nrow(data.focus.pregnant)/nrow(data) * 100, 2)
```

```{r Disaggregation, echo=FALSE, warning=FALSE, message=FALSE}
# Select articles that report no difference in results based on age vs use age as a covariate
data.diff.age <- data %>%
  filter(analysis_no_difference_age == 1)
data.dis.age <- data %>%
  filter(analysis_disaggregation_age == 1)

# Select articles that report no difference in results based on sex vs use sex as a covariate
data.diff.sex <- data %>%
  filter(analysis_no_difference_sex == 1)
data.dis.sex <- data %>%
  filter(analysis_disaggregation_sex == 1)

# Select articles that report no difference in results based on ethnicity vs use ethnicity as a covariate
data.diff.ethnicity <- data %>%
  filter(analysis_no_difference_ethnicity == 1)
data.dis.ethnicity <- data %>%
  filter(analysis_disaggregation_ethnicity == 1)

# Select articles that report no difference in results based on education vs use education as a covariate
data.diff.education <- data %>%
  filter(analysis_no_difference_education == 1)
data.dis.education <- data %>%
  filter(analysis_disaggregation_education == 1)

# Select articles that report no difference in results based on profession vs use profession as a covariate
data.diff.profession <- data %>%
  filter(analysis_no_difference_profession == 1)
data.dis.profession <- data %>%
  filter(analysis_disaggregation_profession == 1)

# Select articles that report no difference in results based on ses vs use ses as a covariate
data.diff.ses <- data %>%
  filter(analysis_no_difference_ses == 1)
data.dis.ses <- data %>%
  filter(analysis_disaggregation_ses == 1)

# Select single studies and disaggregation variables
data.analysis <- data %>%
  select(article_year, analysis_disaggregation_age, analysis_disaggregation_sex, analysis_disaggregation_ethnicity, 
         analysis_disaggregation_education, analysis_disaggregation_profession, analysis_disaggregation_ses)

# Compute whether articles use demographic variables in the analysis
data.analysis <- data.analysis %>%
  mutate(report=if_else(analysis_disaggregation_age == 1 & analysis_disaggregation_sex == 0 & 
                            analysis_disaggregation_ethnicity == 0 & analysis_disaggregation_education == 0 &
                            analysis_disaggregation_profession == 0 & analysis_disaggregation_ses == 0 , "Age", 
                          if_else(analysis_disaggregation_age == 0 & analysis_disaggregation_sex == 1 & 
                            analysis_disaggregation_ethnicity == 0 & analysis_disaggregation_education == 0 &
                            analysis_disaggregation_profession == 0 & analysis_disaggregation_ses == 0 , "Sex",
                          if_else(analysis_disaggregation_age == 1 & analysis_disaggregation_sex == 1 & 
                            analysis_disaggregation_ethnicity == 0 & analysis_disaggregation_education == 0 &
                            analysis_disaggregation_profession == 0 & analysis_disaggregation_ses == 0 , "Age & Sex",
                          if_else(analysis_disaggregation_age == 1 & analysis_disaggregation_sex == 0 & (
                            analysis_disaggregation_ethnicity == 1 | analysis_disaggregation_education == 1 |
                            analysis_disaggregation_profession == 1 | analysis_disaggregation_ses == 1), "Age +1",
                          if_else(analysis_disaggregation_age == 0 & analysis_disaggregation_sex == 1 & (
                            analysis_disaggregation_ethnicity == 1 | analysis_disaggregation_education == 1 |
                            analysis_disaggregation_profession == 1 | analysis_disaggregation_ses == 1), "Sex +1",
                          if_else(analysis_disaggregation_age == 1 & analysis_disaggregation_sex == 1 & (
                            analysis_disaggregation_ethnicity == 1 | analysis_disaggregation_education == 1 |
                            analysis_disaggregation_profession == 1 | analysis_disaggregation_ses == 1), "Age & Sex +1",
                          if_else((analysis_disaggregation_age == 0 | analysis_disaggregation_sex == 0) & (
                            analysis_disaggregation_ethnicity == 1 | analysis_disaggregation_education == 1 |
                            analysis_disaggregation_profession == 1 | analysis_disaggregation_ses == 1), "Other demographics",
                            "No disaggregation", ))))))))

# Count the number of articles for each condition by year of publication
data.analysis <- data.analysis %>%
  count(report, article_year) %>%
  ungroup()

# Compute the proportion of articles per condition by year of publication
data.analysis <- data.analysis %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(report, add=TRUE) %>%
  mutate(per=round(100*n/total,3))
data.analysis$article_year <- as.numeric(data.analysis$article_year)
data.analysis$per <- as.numeric(data.analysis$per)

# Stacked bar chart of the use of demographic variables in the analysis by year of publication
figure12 <- ggplot(data.analysis, aes(x=article_year, y=per)) +
  geom_col(aes(fill=report)) +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  labs(fill="Variable reported") +
  scale_fill_viridis_d() +
  scale_x_continuous(breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(labels=c("0%","25%","50%","75%","100%"), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.title=element_text(face="bold", size=12),
        legend.margin=margin(0,0,0,0), legend.box.margin=margin(-10,0,-10,-10))
```

# Results

### Inclusion & Exclusion

  `r round(percent.included)`% of the sampled articles met eligibility criteria during the review.

```{r, fig.cap="Funding sources and geographical location of the studies. A. Reporting of funding across the years. B. Distribution of articles by study location. The eight most represented countries are individually shown.", warning=FALSE, message=FALSE}
figure1
```

```{r, fig.cap="Funding sources and geographical location of the studies. A. Reporting of funding across the years. B. Distribution of articles by study location. The eight most represented countries are individually shown.", warning=FALSE, message=FALSE}
figure2
```

### Funding

  Funding sources were reported by `r round(percent.funding.name)`% of studies, while funding number was also reported in `r round(percent.funding.num)`% of these cases. Overall, funding by the United States' National Institutes of Health (NIH) represented `r round(percent.funding.NI)`% of the reported funding agencies. `r round(percent.funding.NI.num)`% of the studies funded by the NIH also reported funding number. The second most represented funding agencies were the Australian National Health and Medical Research Council (NHMRC) and the Canadian Institutes of Health Research (CIHR).

```{r, fig.cap="Funding sources and geographical location of the studies. A. Reporting of funding across the years. B. Distribution of articles by study location. The eight most represented countries are individually shown.", warning=FALSE, message=FALSE}
figure3
```

### Geographical location

  `r round(100-percent.country.multiple)`% of articles were conducted in a single country. The geographical location of the study was explicitly reported in `r round(percent.country.reported)`% of studies. The country of study was inferred for the rest of the sampled articles. Inference was primarily based on the first author's affiliation. Overall, `r nrow(country.single)` countries were represented. `r nrow(country.multiple)`% of articles reported multiple countries of study. Figure 2B shows the distribution of study location across time with the eight most represented countries.

```{r, fig.cap="Funding sources and geographical location of the studies. A. Reporting of funding across the years. B. Distribution of articles by study location. The eight most represented countries are individually shown.", warning=FALSE, message=FALSE}
figure4
```

### Sample size

  Sample size was reported in `r round(percent.ss)`% of studies, while `r round(percent.study.single)`% investigated a single sampled population.

```{r, fig.cap="Sample size of the recruited volunteers. Numbers are computed with a log-10 scale.", warning=FALSE, message=FALSE}
figure5
```

### Age

  `r round(percent.age)`% of articles reported a variable describing age. Figure 4A and 4B show the use of various age variables by year and journal of publication. Overall, the mean age of the study populations was `r round(mean(data.age.mean$age_reported_mean_value))` years old.

```{r, fig.cap="To be completed.", warning=FALSE, message=FALSE}
figure6
```
```{r, fig.cap="To be completed.", warning=FALSE, message=FALSE}
figure7
```

```{r, fig.cap="Reporting of age variables. Correlation between the reporting of the mean, standrad deviation of the mean, median, minimum, maximum, interquartile range and other age variables by A. year of publication and B. journal of publication. C. Mean age by year of publication.", message=FALSE, warning=FALSE}
figure8
```

### Sex

  Sex was reported in `r round(percent.sex)`% of the studies. Figure 5 displays the proportion of studies that recruited male subjects, female subjects, both sexes or did not specify the sex of the participants. `r round(percent.sex.male)`% of the studies reporting sex only recruited male participants, while `r round(percent.sex.female)`% only employed females. Out of the studies focusing on a single gender, `r round(percent.sex.male.focus)`% of the male studies focused on a sex dependent feature, while `r round(percent.sex.female.focus)`% of the female studies did. `r round(percent.sex.age)`% of studies reported age by sex.

```{r, fig.cap="Sex inclusion. Proportion of studies that recruited male subjects, female subjects, both sexes, or did not specify the sex of the participants.", warning=FALSE, message=FALSE}
figure9
```

###  Ethnicity, education, profession and socio-economic status

  Other demographics variables were reported in `r round(percent.education)`% of studies for education, `r round(percent.ethnicity)`% for ethnicity, `r round(percent.profession)`% for profession, and `r round(percent.ses)`% for socio-economic status. Figure 6 shows the distribution of this reporting across the years. Figures 7 show the number of categories reported for each variable.

```{r, fig.cap="Reporting of education, ethinicity, profession and socio-economic status.", warning=FALSE, message=FALSE}
figure10
```

```{r, fig.cap="Number of categories reported for education, ethnicity, profession and socio-economic status.", warning=FALSE, message=FALSE}
grid.arrange(figure11A, figure11B, figure11C, figure11D, nrow=2)
```

### Study focus

  `r round(percent.focus.sex)`% of articles focused on a sex dependent feature, while `r round(percent.focus.clinical)`% investigated a clinical feature. `r round(percent.focus.twins)`% of studies focused on twins, `r round(percent.focus.pregnant)`% on pregnant women, `r round(percent.focus.shift)`% on shift workers and `r round(percent.focus.students)`% on university students.

### Analysis disaggregation

  A

```{r, fig.cap="Use of study population characteristics as variables in the analysis."}
figure12
```

# Discussion

In the clinical domain, the need to time therapy based on a patient's individual circadian rhythm has more recently become the focus of the emerging field of chronotherapy or chronotherapeutics [@adam_core_2019; @greco_personalized_2020; @dijk_novel_2020; @hill_optimizing_2020].

The ability to generalise findings from the scientific literature to wide and diverse populations of people hinges upon the representativeness of the study sample with respect to demographic categories. The question to what extent the composition of a given study sample can make the generalisability of findings difficult or impossible has received attention in the field of psychology, where many articles published in prominent journals reflected participants from WEIRD (Western, Educated, Industrialized, Rich, and Democratic) contexts [@henrich_weirdest_2010; @muthukrishna_beyond_2020].

# Conclusion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup